---
title: Methods Workshop in Quantitative Text Analysis Translated in R
author:
  - name: Jisu Kim
    url: https://github.com/jisukimmmm
  - name: Vestin Hategekimana (translator)
    url: https://github.com/CyuHat
date: "21/03/2024"
format: html
---

## Introduction to R language - exercies

### Basic Syntax and Operations:
1. Write a program to calculate the area of a triangle given its base and height.
2. Create a program that converts kilometers per hour to meters per second.
3. Write a Python script that takes a string as input and prints its reverse.

### Conditional Statements and Loops:
1. Create a program that checks whether a given year is a leap year or not.
2. Write a Python script to find the sum of all numbers between 1 and 1000 that are divisible by both 3 and 5.
3. Implement a program to print the first 10 terms of the geometric progression series: 2, 6, 18, 54, ...

### Lists and List Operations:
1. Create a program to find the largest and smallest elements in a list.
2. Write a Python script to find the intersection of two lists.
3. Implement a program to shuffle a deck of cards represented as a list.

### Strings and String Operations:
1. Write a Python script to capitalize the first letter of each word in a sentence.
2. Create a program to find the most frequent character in a given string.
3. Implement a program to check if a given string contains only digits.

### Functions:
1. Create a function to check whether a given number is a perfect square or not.
2. Implement a function to reverse the elements of a list in place.
3. Create a function to calculate the mean (average) of a list of numbers.

### File Handling:
1. Create a program to read a CSV file containing student scores and calculate their average (Use Pandas package).
2. Write a Python script to find and print all lines containing a specific word in a text file.
3. Implement a program to count the number of words in a text file.

### Plotting:
1. Histogram of Student Scores: Create a histogram showing the distribution of student scores.
2. Boxplot of Student Scores: Generate a boxplot to visualize the spread and central tendency of student scores.
3. Scatter Plot of Student Scores: Create a scatter plot to explore the relationship between student scores and student IDs.

1. Calculate the area of a triangle:

```{r}

```

2. Program to convert kilometers per hour to meters per second:

```{r}

```

3. R script to reverse a string:

```{r}

```

### Conditional Statements and Loops:
1. Check if a year is a leap year:

```{r}

```

2. R script to find the sum of all numbers between 1 and 1000 that are divisible by both 3 and 5:

```{r}

```

3. Program to print the first 10 terms of the geometric progression series: 2, 6, 18, 54, ...

```{r}

```

### Lists and List Operations:
1. Find the largest and smallest elements in a list:

```{r}

```

2. R script to find the intersection of two lists:

```{r}

```

3. Program to shuffle a deck of cards represented as a list:

```{r}

```

### Strings and String Operations:
1. Capitalize the first letter of each word in a sentence:

```{r}

```

2. Program to find the most frequent character in a given string:

```{r}

```

3. Program to check if a given string contains only digits:

```{r}

```

### Functions:
1. Check if a number is a perfect square:

```{r}

```

2. Implement a function to reverse the elements of a list in place:

```{r}

```

3. Create a function to calculate the mean (average) of a list of numbers:

```{r}

```

### File Handling:
1. Read a CSV file containing student scores and calculate their average (Use Pandas package):

```{r}

```

2. Write an R script to find and print all lines containing a specific word in a text file:

```{r}

```

3. Implement a program to count the number of words in a text file:

```{r}

```

### Plotting:
1. Histogram of Student Scores:

```{r}

```

2. Boxplot of Student Scores:

```{r}

```

3. Scatter Plot of Student Scores:

```{r}

```